```{r intro-to-dtplyr, include = FALSE}
eval_dtplyr <- FALSE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_dtplyr <- Sys.getenv("GLOBAL_EVAL")
```


```{r, eval = eval_dtplyr, include = FALSE}
library(data.table)
library(dtplyr)
library(dplyr)
library(lobstr)
library(fs)
library(purrr)
```


data.table is a high performance version of data.frame
`fread` is a faster file read. Vroom is faster, but remember, it is not putting everything into memory.
concise syntax DT[i,j,by] # kind of like javascript. It is supposed to be different.



# Introduction to `dtplyr`

## `dtplyr` basics
*Load data into R via `data.table`, and then wrap it with `dtplyr`*

dtyplyr automatically translates data.table to be used with dplyr grammar.
Lazy evaluation: only grabs a lazy evaluation first. 

1. Load the `data.table`, `dplyr`, `dtplyr`, `purrr` and `fs` libraries
    ```{r, eval = eval_dtplyr}
    library(data.table)
    library(dplyr)
    library(dtplyr)
    library(purrr)
    library(fs)
    ```

2. Read the **transactions.csv** file, from the **/usr/share/class/files** folder. Use the `fread()` function to load the data into a variable called `transactions`
    ```{r, eval = eval_dtplyr}
    transactions <- dir_ls("/usr/share/class/files", glob = "*.csv") %>%
       map(fread) %>%
       rbindlist()
    
    
    # map from purrr.
    # data.table needs extra work by using map.
    # for each dir_ls item, fread it into a data.table then rbindlist it (take resulting list and bind each lists)
    ```

3. Preview the data using `glimpse()`
    ```{r, eval = eval_dtplyr}
    glimpse(transactions)
    ```

4. Use `lazy_dt()` to "wrap" the `transactions` variable into a new variable called `dt_transactions`
    ```{r, eval = eval_dtplyr}
    # lazily evaluated data table ..
    (dt_transactions <- lazy_dt(transactions))
    ```

5. View `dt_transactions` structure with `glimpse()`
    ```{r, eval = eval_dtplyr}
    glimpse(dt_transactions)
    ```

## Object sizes
*Confirm that `dtplyr` is not making copies of the original `data.table`*

1. Load the `lobstr` library
    ```{r, eval = eval_dtplyr}
    library(lobstr)
    # looks at data size.
    ```

2. Use `obj_size()` to obtain `transactions`'s size in memory
    ```{r, eval = eval_dtplyr}
    obj_size(transactions)
    ```

3. Use `obj_size()` to obtain `dt_transactions`'s size in memory
    ```{r, eval = eval_dtplyr}
    obj_size(dt_transactions)
    ```

4. Use `obj_size()` to obtain `dt_transactions` and `transactions` size in memory together
    ```{r, eval = eval_dtplyr}
    obj_size(transactions,dt_transactions) # memory together is occupying the amount of dt_transactions. dt_transactions is only referencing the original transactions without creating extra overhead.
    ```

## How `dtplyr` works
*Under the hood view of how `dtplyr` operates `data.table` objects*

avoids copying things as much as possible.
dtyplyr by default does not update. Protects them from accidentally updating by reference.

`lazy_dt()` has option that says `immutable = FALSE` to prevent dplyr from making copies.

1. Use `dplyr` verbs on top of `dt_transactions` to obtain the total sales by month
    ```{r, eval = eval_dtplyr}
    
    # dt_transaction is a lazily evaluated data table.
    # we are summarizing sales using the lazy dt.
    dt_transactions %>%
      group_by(date_month) %>%
      summarise(total_sales = sum(price)) # %>% collect() # to collect all values.
    
    # `_DT1`[, .(total_sales = sum(price)), keyby = .(date_month)]
    # is what you would do in data.table to get this done.
    # only executes on summary of the data. FAST! 
    ```

2. Load the above code into a variable called `by_month`
    ```{r, eval = eval_dtplyr}
      by_month <- dt_transactions %>%
      group_by(date_month) %>%
      summarise(total_sales = sum(price))
    ```

3. Use `show_query()` to see the `data.table` code that `by_month` actually runs
    ```{r, eval = eval_dtplyr}
    show_query(by_month) # most efficient way to use data.table code.
    ```

4. Use `glimpse()` to view how `by_month`, instead of modifying the data, only adds steps that will later be executed by `data.table`
    ```{r, eval = eval_dtplyr}
    glimpse(by_month)
    # contains reference to data.table
    # also instructions for when this object needs to really be executed.
    ```
    
5. Create a new column using `mutate()`
    ```{r, eval = eval_dtplyr}
    # using mutate 
    dt_transactions %>%
      mutate(new_field = log(price+.01))
    ```

6. Use `show_query()` to see the `copy()` command being used
    ```{r, eval = eval_dtplyr}
    # showing that it is doing it on a copy. Not necessarily the original object.
    # This is the default behavior
        dt_transactions %>%
      mutate(new_field = log(price+.01)) %>% 
      show_query()
    ```

7. Check to confirm that the new column *did not* persist in `dt_transactions`
    ```{r, eval = eval_dtplyr}
    dt_transactions
    ```

8. Use `lazy_dt()` with the `immutable` argument set to `FALSE` to avoid the copy
    ```{r, eval = eval_dtplyr}
    transactions2 <- copy(transactions) # transactions is the beginning file that is read in.
    # by copying, we are actually creating distinct objects.
    
    m_transactions <- lazy_dt(transactions2, immutable = FALSE) # to avoid the copy, False. 
    # By default it is true.
    obj_size(transactions,transactions2)
    # see here that it has made a distinct object because size has doubled.
    
    
    # When you False, then we don't have that copy command.
    # When you True, then we have copy command.
    
    # Copy will make sure that it doesn't override.
    
    # immutable = FALSE will allow for full-benefit of data.table.
    ```

Remember, the default way data.table works is it updates whenever the query is made. dtplyr allows for more R-like grammar.


    ```{r, eval = eval_dtplyr}
    m_transactions
    ```

9. Create a `new_field` column in `m_transactions` using `mutate()`
    ```{r, eval = eval_dtplyr}
    m_transactions %>% 
      mutate(new_field = log(price + .01))
    ```

10. Use `show_query()` to see that `copy()` is no longer being used
    ```{r, eval = eval_dtplyr}
    m_transactions %>% show_query()
    ```

11. Inspect `m_transactions` to see that `new_field` has persisted
    ```{r, eval = eval_dtplyr}
    m_transactions
    ```

## Working with `dtplyr`
*Learn data conversion and basic visualization techniques*

1. Use `as_tibble()` to convert the results of `by_month` into a `tibble`
    ```{r, eval = eval_dtplyr}
    by_month %>%
      as_tibble()
    # we can have it as tibble if we want to interact the way we usually know.
    ```

2. Load the `ggplot2` library
    ```{r, eval = eval_dtplyr}
    library(ggplot2)
    ```

3. Use `as_tibble()` to convert before creating a line plot 
    ```{r, eval = eval_dtplyr}
    by_month %>%
      as_tibble() %>%
      ggplot() +
      geom_line(aes(date_month, total_sales))
    ```

## Pivot data
*Review a simple way to aggregate data faster, and then pivot it as a tibble*

1. Load the `tidyr` library
    ```{r, eval = eval_dtplyr}
    library(tidyr)
    ```

2. Group `db_transactions` by `date_month` and `date_day`, then aggregate `price` into `total_sales`
    ```{r, eval = eval_dtplyr}
    dt_transactions %>%
      group_by(date_month, date_day) %>% 
      summarise(total_sales = sum(price))
    ```

3. Copy the aggregation code above, **collect it into a `tibble`**, and then use `pivot_wider()` to make the `date_day` the column headers.
    ```{r, eval = eval_dtplyr}
    dt_transactions %>%
      group_by(date_month, date_day) %>% 
      summarise(total_sales = sum(price)) %>%
      as_tibble() %>% # Take advantage of lazy_dt. When you are ready to use it with ggplot, collect it as tibble for the moment.
      pivot_wider(names_from = date_day, values_from = total_sales) 
    # fast way to pivot using data.table queries
    
    ```
