```{r db-visualizations, include = FALSE}
eval_viz <- FALSE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_viz <- Sys.getenv("GLOBAL_EVAL")
```

```{r, eval = eval_viz, include = FALSE}
library(connections)
library(dplyr)
library(dbplyr)
library(dbplot)
library(ggplot2)
library(leaflet)
library(config)
```


# Data Visualizations

## Simple plot
*Practice pushing the calculations to the database*

1. Load the `connections`, `dplyr`, `dbplyr`, and `config` libraries
    ```{r, eval = eval_viz}
    library(connections)
    library(dplyr)
    library(dbplyr)
    library(config)
    ```

2. Use `connection_open()` to open a Database connection
    ```{r, eval = eval_viz}
    con <- connection_open(
      RPostgres::Postgres(),
      host =  "localhost",
      user = get("user", config = "dev"),
      password = get("pwd", config = "dev"),
      port = 5432,
      dbname = "postgres",
      bigint = "integer"
    )
    ```

3. Use `tbl()` to create a pointer to the **v_orders** table
    ```{r, eval = eval_viz}
    orders <- tbl(con, in_schema("retail","v_orders"))
    class(orders)
    ```

4. Use `collect()` bring back the aggregated results into a "pass-through" variable called `by_year`
    ```{r, eval = eval_viz}
    by_year <- orders %>%
      count(date_year) %>%
      collect() # bring results of this query into R
    ```

5. Preview the `by_year` variable
    ```{r, eval = eval_viz}
    class(by_year)
    ```

6. Load the `ggplot2` library
    ```{r, eval = eval_viz}
    library(ggplot2)
    ```


7. Plot results using `ggplot2`
    ```{r, eval = eval_viz}
    ggplot(by_year) +
      geom_col(aes(date_year, n)) 
    ```

8. Using the code in this section, create a single piped code set which also creates the plot
    ```{r, eval = eval_viz}
    orders %>%
  count(date_year) %>%
  collect() %>%
  ggplot() +            # < Don't forget to switch to `+`
  geom_col(aes(date_year, n)) 
    ```

## Plot in one code segment
*Practice going from `dplyr` to `ggplot2` without using pass-through variable, great for EDA*

1. Summarize the order totals in a new variable called `sales`
    ```{r, eval = eval_viz}
    orders %>%
  summarise(sales = sum(order_total))
    ```

2. Summarize the order totals grouped by `date_year` in a new variable called `sales`
    ```{r, eval = eval_viz}
    orders %>%
      group_by(date_year) %>%
      summarise(sales = sum(order_total))
    ```
    
3. Summarize the order totals grouped by `date_year` in a new variable called `sales` and plot the results
    ```{r, eval = eval_viz}
    orders %>%
  group_by(date_year) %>%
  summarise(sales = sum(order_total)) %>%
  ggplot() +
  geom_col(aes(date_year, sales))
    ```

4. Switch the calculation to reflect the average of the order sale total
    ```{r, eval = eval_viz}
    orders %>%
  group_by(date_year) %>%
  summarise(sales = mean(order_total)) %>%
  ggplot() +
  geom_col(aes(date_year, sales))
    ```

## Create a histogram
*Use the `dbplot` package to easily create a histogram*

1. Load the `dbplot` package
    ```{r, eval = eval_viz}
    library(dbplot)
    ```


Showing the query for creating new bins that can be returned by histogram!
```{r}
orders %>% 
  mutate(bin = !!db_bin(order_total,binwidth = 10)) %>% 
  show_query()
```


2. Use the `dbplot_histogram()` to build the histogram
    ```{r, eval = eval_viz}
    orders %>%
      dbplot_histogram(order_total)
    ```

Really good because we don't have to directly pull data and bring it to local memory. We can just directly query it!

3. Adjust the `binwidth` to 10
    ```{r, eval = eval_viz}
    orders %>%
  dbplot_histogram(order_total, binwidth = 10) + theme_bw()
    ```

## Raster plot
*Use `dbplot`'s raster graph*

Usually, trends and scatterplots are just way too big. 
Instead of trying to create a scatterplot that creates millions of records.

It shows the density of points accross variables. 

1. Use a `dbplot_raster()` to visualize `order_qty` versus `order_total`
    ```{r, eval = eval_viz}
    orders %>%
      dbplot_raster(order_qty, order_total)
    ```

The colors are a heat map for density.
Plotting up an aggregation.


2. Change the plot's resolution to 10
    ```{r, eval = eval_viz}
        orders %>%
      dbplot_raster(order_qty, order_total, resolution = 10)
    ```

## Using the `compute` functions

Combination of both approaches!! 
Histogram & raster.
Utilize summarization in density form.
Summarize, return results, aggregate, end.

We probably won't want this behavior with map data like this..
```{r}
orders %>% dbplot_raster(customer_lon,customer_lat, resolution = 10)
```


1. Instead use the `db_compute_raster2()` function to get underlying results that feed the plot
    ```{r, eval = eval_viz}
    locations <- orders %>%
      db_compute_raster2(customer_lon, customer_lat, resolution = 10)
    ```

What's the difference between compute_raster and compute_raster2
It gives me both boundaries for longditude and lattitude.

2. Preview the `locations` variable
    ```{r, eval = eval_viz}
    locations
    ```

3. Load the `leaflet` library
    ```{r, eval = eval_viz}
    library(leaflet)
    ```

4. Pipe `location` into the `leaflet()` function, and then pipe that into the `addTiles()` function
    ```{r, eval = eval_viz}
    locations %>%
      leaflet() %>% 
        addTiles() 
    ```

5. Add the `addRectangles()` function using the longitude and latitude variables 
```{r, eval = eval_viz}
locations %>%
  leaflet() %>% 
    addTiles() %>%
    addRectangles(
      ~customer_lon, 
      ~customer_lat, 
      ~customer_lon_2,
      ~customer_lat_2
    )
```

6. Add the `fillOpacity` argument to the `addRectangles()` step, use ``n()`` as the value for it
```{r, eval = eval_viz}
locations %>%
  leaflet() %>% 
    addTiles() %>%
    addRectangles(
      ~customer_lon, 
      ~customer_lat, 
      ~customer_lon_2,
      ~customer_lat_2,
      fillOpacity = ~`n()`
    )
```

7. Modify `fillOpacity` to be calculated as a percentage against the maximum number of orders
```{r, eval = eval_viz}
locations %>%
  leaflet() %>% 
    addTiles() %>%
    addRectangles(
      ~customer_lon, 
      ~customer_lat, 
      ~customer_lon_2,
      ~customer_lat_2,
      fillOpacity = ~(`n()` / max(`n()`))
    )
```

8. Add the `popup` argument with the following instruction as its value: `~paste0("<p>No of orders: ",  `n()`,"</p>")`
```{r, eval = eval_viz}
locations %>%
  leaflet() %>% 
    addTiles() %>%
    addRectangles(
      ~customer_lon, 
      ~customer_lat, 
      ~customer_lon_2,
      ~customer_lat_2,
      fillOpacity = ~(`n()` / max(`n()`)),
      popup = ~paste0("<p>No of orders: ",  `n()`,"</p>")
    )
```

If I have other information such as area name I can also include it in the popup arguments.


9. Disconnect from the database using `connection_close`
    ```{r}
    connection_close(con)
    ```

